Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2017-02-05 nips2016/nips2016.tex
rename folder: images -> figures
nips2016 final paper revision

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[english]{babel}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{xfrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
%\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{multirow}
%\usepackage{svg}
\usepackage{setspace}
\usepackage{color}
\usepackage{capt-of}

% Use less space
\usepackage{titlesec}		% space before / after headings
\titlespacing\section{0pt}{5pt plus 1pt minus 2pt}{1.8pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{5pt plus 1pt minus 2pt}{2.7pt plus 1pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}

@@

\usepackage{parskip} \setlength{\parskip}{2.1pt} \setlength{\parindent}{1em}

@@

\author{
	Michaël Defferrard\affe \hspace{1cm} Xavier Bresson\affs \hspace{1cm} Pierre Vandergheynst\affe %\thanks{further info} \\
	%Department of Electrical Engineering \\
	%École Polytechnique Fédérale de Lausanne (EPFL) \\
	\vspace{0.2cm} \\
	\affe EPFL, Lausanne, Switzerland \hspace{1cm} \affs Nanyang Technological University, Singapore \\
	\texttt{\{michael.defferrard,pierre.vandergheynst\}@epfl.ch \hfill xavier.bresson@ntu.edu.sg} \\
}

@@

User data on social networks, gene data on biological regulatory networks, log
data on telecommunication networks, or text documents on word embeddings are
important examples of data lying on irregular or non-Euclidean domains which
can be structured with graphs, which are universal representations of
heterogeneous pairwise relationships. Graphs can encode complex geometric
structures and can be studied with strong mathematical tools such as spectral
graph theory \cite{book:Chung97Spectral}.

A generalization of CNNs to graphs is not straightforward as the convolution
and pooling operators are only defined for regular grids. This makes this
extension challenging, both theoretically and implementation-wise. The major
bottleneck of generalizing CNNs to graphs, and one of the primary goals of this
work, is the definition of localized graph filters which are efficient to
evaluate and learn.
Precisely, the main contributions of this work are summarized below.
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A spectral graph theoretical formulation
	of CNNs on graphs built on established tools in GSP
	\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
\item \textbf{Strictly localized filters.} Enhancing
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, the proposed spectral filters
	are provable to be strictly localized in a ball of radius $K$, i.e. $K$
	hops from the central vertex.  
\item \textbf{Low computational complexity.} The evaluation complexity of our
	filters is linear w.r.t. the filters support's size $K$ and the number of
	edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we
	have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor
	(NN) graphs, leading to a linear complexity w.r.t the input data size.
	Moreover, this method avoids the Fourier basis altogether, thus the
	expensive eigenvalue decomposition (EVD) necessary to compute it as well as
	the need to store the basis, a matrix of size $n^2$. That is especially
	relevant when working with limited GPU memory. Besides the data, our method
	only requires to store the Laplacian, a sparse matrix of $|\E|$ non-zero
	values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices as a binary tree
	structure, is analog to pooling of 1D signals.
\item \textbf{Experimental results.} We present multiple experiments that
	ultimately show that our formulation is (i) a useful model, (ii)
	computationally efficient and (iii) superior both in accuracy and
	complexity to the pioneer spectral graph CNN introduced in
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. We also show that our graph
	formulation performs similarly as a classical CNNs on MNIST and study the
	impact of various graph constructions on classification performance. The
	TensorFlow \cite{abadi_tensorflow_2016} code to reproduce our results and
	apply the model to other data is available as an open-source
	software.\footnote{\url{https://github.com/mdeff/cnn_graph}}
\end{enumerate}

\begin{figure}[t]
\centering
\includegraphics[width=0.87\textwidth]{images/illustrationCNNgraphs}
\caption{Architecture of the proposed CNN on graphs.
	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
}
\label{fig_illustration}
\end{figure}

@@

There are two strategies to define convolutional filters; either from a spatial
approach or from a spectral approach. By construction, spatial approaches
provide filter localization via the finite size of the kernel. However,
although graph convolution directly in the spatial domain is conceivable, it
also faces the challenge of matching local neighborhoods, as pointed out in
\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique
mathematical definition of translation on graphs from a spatial perspective. On
the other side, a spectral approach provides a well-defined translation
operator on graphs via convolutions with a Kronecker delta implemented in the
spectral domain \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
The convolution theorem \cite{book:Mallat99wavelets} defines convolutions as
linear operators that diagonalize in the Fourier basis (represented by the
eigenvectors of the Laplacian operator). However, a filter defined in the
spectral domain is not naturally localized and translations are costly due to
the $\bO(n^2)$ multiplication with the graph Fourier basis. Both limitations
can however be overcome with a special choice of filter parametrization.

@@

\paragraph{Graph Fourier Transform.} We are interested in processing signals
defined on undirected and connected graphs $\G=(\V,\E,W)$, where $\V$ is a
finite set of vertices with $|\V|=n$, $\E$ is a set of edges and $W$ is a
weighted adjacency matrix encoding the connection weight between two vertices. A
signal $x: \V \rightarrow \R$ defined on the nodes of the graph may be regarded
as a vector $x \in \R^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
An essential operator in spectral graph analysis is the graph Laplacian
\cite{book:Chung97Spectral}, which combinatorial definition is $L = D - W \in
\R^{n \times n}$ where $D \in \R^{n \times n}$ is the diagonal degree matrix
with $D_{ii} = \sum_j W_{ij}$, and normalized definition is $L = I_n -
D^{-1/2} W D^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real
symmetric positive semidefinite matrix, it has a complete set of orthonormal
eigenvectors $\{u_l\}_{l=0}^{n-1} \in \R^n$, known as the graph Fourier modes,
and their associated ordered real nonnegative eigenvalues
$\{\lambda_l\}_{l=0}^{n-1}$, identified as the frequencies of the graph. The
Laplacian is indeed diagonalized by the Fourier basis $U=[u_0, \ldots, u_{n-1}]
\in \R^{n \times n}$ such that $L = U \Lambda U^T$ where $\Lambda =
\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$. We can now
define the graph Fourier transform (GFT) of a spatial signal $x \in \R^n$ as
$\hat{x} = U^T x \in \R^n$, and its inverse as $x = U \hat{x}$
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Similarly to the
Fourier transform on Euclidean spaces, the GFT enables the formulation of
fundamental operations such as filtering.

@@

\paragraph{Polynomial parametrization for localized filters.} There are however
two limitations with non-parametric filters: (i) they are not localized in
space and (ii) their learning complexity is in $\bO(n)$, the dimensionality of
the data. These issues can be overcome with the use of a polynomial filter
% a Laplacian-based polynomial spectral filter
\begin{equation} \label{eq:filt_poly}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is
given by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
delta function $\delta_i \in \R^n$. By \cite[Lemma
5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum
number of edges connecting two vertices on the graph.  Consequently, spectral
filters represented by $K^\text{th}$-order polynomials of the Laplacian are
exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the
support size of the filter, and thus the same complexity as classical CNNs.

@@

Recall that the Chebyshev polynomial $T_k(x)$ of order $k$ may be generated by
the stable recurrence relation $T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)$ with $T_0 =
1$ and $T_1 = x$. These polynomials form an orthogonal basis for $L^2([-1,1], dy
/ \sqrt{1-y^2})$, the Hilbert space of square integrable functions with respect
to the measure $dy/\sqrt{1-y^2}$. A filter can thus be parametrized as a
truncated expansion of order $K-1$ such that
\begin{equation} \label{eq:filt_cheby}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}),
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of Chebyshev coefficients and
$T_k(\tilde{\Lambda}) \in \R^{n \times n}$ is the Chebyshev polynomial of order
$k$ evaluated at $\tilde{\Lambda} = 2 \Lambda / \lambda_{max} - I_n$, a diagonal
matrix of scaled eigenvalues that lie in $[-1,1]$.
The filtering operation can then be written as $y = g_\theta(L) x
= \sum_{k=0}^{K-1} \theta_k T_k(\tilde{L}) x$, where $T_k(\tilde{L}) \in \R^{n
\times n}$ is the Chebyshev polynomial of order $k$ evaluated at the scaled
Laplacian $\tilde{L} = 2 L / \lambda_{max} - I_n$. %Note that the spectrum of the
%normalized Laplacian is bounded by $2$ \cite{book:Chung97Spectral}, such that
%the scaling can simply be $\tilde{L} = L - I_n$. 
Denoting $\bar{x}_k =
T_k(\tilde{L})x \in \R^n$, we can use the recurrence relation to compute
$\bar{x}_k = 2\tilde{L} \bar{x}_{k-1} - \bar{x}_{k-2}$ with $\bar{x}_0 = x$ and
$\bar{x}_1 = \tilde{L}x$. The entire filtering operation
	$y = g_\theta(L) x = [\bar{x}_0, \ldots, \bar{x}_{K-1}] \theta$
then costs $\bO(K|\E|)$ operations.

@@

Graclus \cite{art:DhillonGuanKulis07Graclus}, built on Metis
\cite{art:KarypisKumar98Metis}, uses a greedy algorithm to compute successive
coarser versions of a given graph and is able to minimize several popular
spectral clustering objectives. We chose the normalized cut
\cite{art:ShiMalik00NCut}, which is an excellent clustering energy. Graclus'
greedy rule consists, at any given coarsening level, in picking an unmarked
vertex $i$ and matching it with one of its unmarked neighbors $j$ that
maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$.  The two matched
vertices are then marked and the coarsened weights are set as the sum of their
weights. The matching is repeated until all nodes have been explored. This is
an extremely fast coarsening scheme which divides the number of nodes by
approximately two (there may exist a few singletons, non-matched nodes) from
one level to the next coarser level.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{images/coarsening}
\caption{\textbf{Example of Graph Coarsening and Pooling.} Let us carry out a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
\R^8$ living on $\G_0$, the finest graph given as input. Note that it originally
possesses $n_0 = |\V_0| = 8$ vertices, arbitrarily ordered. For a pooling of
size 4, two coarsenings of size 2 are needed: Graclus first gives $\G_1$ of size
$n_1 = |\V_1| = 5$, then $\G_2$ of size $n_2 = |\V_2| = 3$, the coarsest graph.
Sizes are thus set to $n_2 = 3$, $n_1 = 6$, $n_0 = 12$ and fake nodes (in blue)
are added to $\V_1$ (1 node) and $\V_0$ (4 nodes) to pair with the singeltons
(in orange), such that each node has exactly two children.  Nodes in $\V_2$ are
then arbitrarily ordered and nodes in $\V_1$ and $\V_0$ are ordered
consequently.  At that point the arrangement of vertices in $\V_0$ permits a
regular 1D pooling on $x \in \R^{12}$ such that $z = [\max(x_0,x_1),
\max(x_4,x_5,x_6), \max(x_8,x_9,x_{10})] \in \R^3$, where the signal components
$x_2,x_3,x_7,x_{11}$ are set to a neutral value.}
\label{fig:pooling}
\end{figure}

Pooling operations are carried out many times and must be efficient. After
coarsening, the vertices of the input graph and its coarsened versions are not
arranged in any meaningful way. Hence, a direct application of the pooling
operation would need a table to store all matched vertices. That would result
in a memory inefficient, slow, and hardly parallelizable implementation. It is
however possible to structure the arrangement of the vertices such that a graph
pooling operation becomes as efficient as a 1D pooling. We proceed in two
steps: (i) create a balanced binary tree and (ii) rearrange the vertices.
After coarsening, each node has either two parents, if it was matched at the
finer level, or one, if it was not, i.e the node was a singleton. From the
coarsest to finest level, fake nodes, i.e.  disconnected nodes, are added to
pair with the singletons such that each node has two children. This structure
is a balanced binary tree: (i) regular nodes (and singletons) either have two
regular nodes (e.g. level 1 vertex 0 in \figref{pooling}) or (ii) one singleton
and a fake node as children (e.g. level 2 vertex 0 in \figref{pooling}), and
(iii) fake nodes always have two fake nodes as children (e.g. level 1 vertex 1
in \figref{pooling}). Input signals are initialized with a neutral value at the
fake nodes, e.g. 0 when using a ReLU activation with max pooling. Because these
nodes are disconnected, filtering does not impact the initial neutral value.
While those fake nodes do artificially increase the dimensionality thus the
computational cost, we found that, in practice, the number of singletons left
by Graclus is quite low.
% < 10%? compared to the number of vertices.
Arbitrarily ordering the nodes at the coarsest level, then propagating this
ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
children, produces a regular ordering in the finest level. Regular in the sense
that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
rearranged graph signal is analog to pooling a regular 1D signal.
\figref{pooling} shows an example of the whole coarsening and pooling process.
This regular arrangement makes the pooling operation very efficient and
satisfies parallel architectures such as GPUs as memory accesses are local,
i.e.  matched nodes do not have to be fetched.

@@

The works of \cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF} introduced the
concept of constructing a local receptive field to reduce the number of learned
parameters. The idea is to group together features based upon a measure of
similarity such as to select a limited number of connections between two
successive layers. While this model reduces the number of parameters by
exploiting the locality assumption, it did not attempt to exploit any
stationarity property, i.e. no weight-sharing strategy.
%This approach is however not able to extract similar groups over the data
%domain, as it is not clear how to define convolutional filters on these
%groups.
The authors of \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} used this idea for
their spatial formulation of graph CNNs. They use the weighted graph to define
the local neighborhood and compute a multiscale clustering of the graph for the
pooling operation. Inducing weight sharing in a spatial construction is however
challenging, as it requires to select and order the neighborhoods when a
problem-specific ordering (spatial, temporal, or otherwise) is missing.

A spatial generalization of CNNs to 3D-meshes, a class of smooth
low-dimensional non-Euclidean spaces, was proposed in \cite{masci2015geodesic}.
The authors used geodesic polar coordinates to define convolution operations on
mesh patches, and formulated a deep learning architecture which allows
comparison across different manifolds. They obtained state-of-the-art results
for 3D shape recognition tasks.

@@

To validate our model, we applied it to the Euclidean case on the benchmark
MNIST classification problem \cite{pro:LeCunBottouBengioHaffner98MNIST}, a
dataset of 70,000 digits represented on a 2D grid of size $28 \times 28$. For
our graph model, we construct a $8$-NN graph of the 2D grid which produces a
graph of $n = |\V| = 976$ nodes ($28^2 = 784$ pixels and 192 fake nodes as
explained in \secref{pooling}) and $|\E| = 3198$ edges. Following standard
practice, the weights of a $k$-NN similarity graph (between features) are
computed as
%$W_{ij} = \exp(-\|z_i - z_j\|_2^2 / \sigma^2)$
\begin{equation} \label{eq:knngraph}
	W_{ij} = \exp \left( -\frac{\|z_i - z_j\|_2^2}{\sigma^2} \right)
\end{equation}
where $z_i$ is the 2D coordinate of pixel $i$.
% and \sigma is set using the scheme proposed in

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[english]{babel}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{xfrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
\graphicspath{{figures/}}
%\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{multirow}
%\usepackage{svg}
\usepackage{setspace}
\usepackage{color}
\usepackage{capt-of}

% Use less space
\usepackage{titlesec}		% space before / after headings
\titlespacing\section{0pt}{3.3pt plus 1pt minus 2pt}{1.8pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{3.3pt plus 1pt minus 2pt}{2.7pt plus 1pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}

@@

\usepackage{parskip} \setlength{\parskip}{3.9pt} %\setlength{\parindent}{1em}

@@

\author{
	Michaël Defferrard \hspace{1cm} Xavier Bresson \hspace{1cm} Pierre Vandergheynst %\thanks{further info} \\
	%Department of Electrical Engineering \\
	%École Polytechnique Fédérale de Lausanne (EPFL) \\
	\vspace{0.2cm} \\
	%\affe EPFL, Lausanne, Switzerland \hspace{1cm} \affs Nanyang Technological University, Singapore \\
	EPFL, Lausanne, Switzerland \\
	\texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} \\
}

@@

User data on social networks, gene data on biological regulatory networks, log
data on telecommunication networks, or text documents on word embeddings are
important examples of data lying on irregular or non-Euclidean domains that
can be structured with graphs, which are universal representations of
heterogeneous pairwise relationships. Graphs can encode complex geometric
structures and can be studied with strong mathematical tools such as spectral
graph theory \cite{book:Chung97Spectral}.

A generalization of CNNs to graphs is not straightforward as the convolution
and pooling operators are only defined for regular grids. This makes this
extension challenging, both theoretically and implementation-wise. The major
bottleneck of generalizing CNNs to graphs, and one of the primary goals of this
work, is the definition of localized graph filters which are efficient to
evaluate and learn.
Precisely, the main contributions of this work are summarized below.
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A spectral graph theoretical formulation
	of CNNs on graphs built on established tools in graph signal processing (GSP).
	\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
\item \textbf{Strictly localized filters.} Enhancing
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, the proposed spectral filters
	are provable to be strictly localized in a ball of radius $K$, i.e. $K$
	hops from the central vertex.  
\item \textbf{Low computational complexity.} The evaluation complexity of our
	filters is linear w.r.t. the filters support's size $K$ and the number of
	edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we
	have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor
	(NN) graphs, leading to a linear complexity w.r.t the input data size $n$.
	Moreover, this method avoids the Fourier basis altogether, thus the
	expensive eigenvalue decomposition (EVD) necessary to compute it as well as
	the need to store the basis, a matrix of size $n^2$. That is especially
	relevant when working with limited GPU memory. Besides the data, our method
	only requires to store the Laplacian, a sparse matrix of $|\E|$ non-zero
	values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices as a binary tree
	structure, is analog to pooling of 1D signals.
\item \textbf{Experimental results.} We present multiple experiments that
	ultimately show that our formulation is (i) a useful model, (ii)
	computationally efficient and (iii) superior both in accuracy and
	complexity to the pioneer spectral graph CNN introduced in
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. We also show that our graph
	formulation performs similarly to a classical CNNs on MNIST and study the
	impact of various graph constructions on performance. The
	TensorFlow \cite{abadi_tensorflow_2016} code to reproduce our results and
	apply the model to other data is available as an open-source
	software.\footnote{\url{https://github.com/mdeff/cnn_graph}}
\end{enumerate}

\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{architecture}
	\caption{Architecture of a CNN on graphs and the four ingredients of a (graph) convolutional layer.}
	\label{fig:architecture}
\end{figure}

@@

There are two strategies to define convolutional filters; either from a spatial
approach or from a spectral approach. By construction, spatial approaches
provide filter localization via the finite size of the kernel. However,
although graph convolution in the spatial domain is conceivable, it
faces the challenge of matching local neighborhoods, as pointed out in
\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique
mathematical definition of translation on graphs from a spatial perspective. On
the other side, a spectral approach provides a well-defined localization
operator on graphs via convolutions with a Kronecker delta implemented in the
spectral domain \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
The convolution theorem \cite{book:Mallat99wavelets} defines convolutions as
linear operators that diagonalize in the Fourier basis (represented by the
eigenvectors of the Laplacian operator). However, a filter defined in the
spectral domain is not naturally localized and translations are costly due to
the $\bO(n^2)$ multiplication with the graph Fourier basis. Both limitations
can however be overcome with a special choice of filter parametrization.

@@

\paragraph{Graph Fourier Transform.} We are interested in processing signals
defined on undirected and connected graphs $\G=(\V,\E,W)$, where $\V$ is a
finite set of $|\V|=n$ vertices, $\E$ is a set of edges and $W \in \R^{n \times n}$ is a
weighted adjacency matrix encoding the connection weight between two vertices. A
signal $x: \V \rightarrow \R$ defined on the nodes of the graph may be regarded
as a vector $x \in \R^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
An essential operator in spectral graph analysis is the graph Laplacian
\cite{book:Chung97Spectral}, which combinatorial definition is $L = D - W \in
\R^{n \times n}$ where $D \in \R^{n \times n}$ is the diagonal degree matrix
with $D_{ii} = \sum_j W_{ij}$, and normalized definition is $L = I_n -
D^{-1/2} W D^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real
symmetric positive semidefinite matrix, it has a complete set of orthonormal
eigenvectors $\{u_l\}_{l=0}^{n-1} \in \R^n$, known as the graph Fourier modes,
and their associated ordered real nonnegative eigenvalues
$\{\lambda_l\}_{l=0}^{n-1}$, identified as the frequencies of the graph. The
Laplacian is indeed diagonalized by the Fourier basis $U=[u_0, \ldots, u_{n-1}]
\in \R^{n \times n}$ such that $L = U \Lambda U^T$ where $\Lambda =
\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$.
The graph Fourier transform of a signal $x \in \R^n$ is then defined as
$\hat{x} = U^T x \in \R^n$, and its inverse as $x = U \hat{x}$
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. As on Euclidean
spaces, that transform enables the formulation of fundamental
operations such as filtering.

@@

\paragraph{Polynomial parametrization for localized filters.} There are however
two limitations with non-parametric filters: (i) they are not localized in
space and (ii) their learning complexity is in $\bO(n)$, the dimensionality of
the data. These issues can be overcome with the use of a polynomial filter
% a Laplacian-based polynomial spectral filter
\begin{equation} \label{eq:filt_poly}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is
given by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
(L^k)_{i,j}$, where the kernel is localized via a convolution with a Kronecker
delta function $\delta_i \in \R^n$. By \cite[Lemma
5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum
number of edges connecting two vertices on the graph.  Consequently, spectral
filters represented by $K^\text{th}$-order polynomials of the Laplacian are
exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the
support size of the filter, and thus the same complexity as classical CNNs.

@@

Recall that the Chebyshev polynomial $T_k(x)$ of order $k$ may be computed by
the stable recurrence relation $T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)$ with $T_0 =
1$ and $T_1 = x$. These polynomials form an orthogonal basis for $L^2([-1,1], dy
/ \sqrt{1-y^2})$, the Hilbert space of square integrable functions with respect
to the measure $dy/\sqrt{1-y^2}$. A filter can thus be parametrized as the
truncated expansion
\begin{equation} \label{eq:filt_cheby}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}),
\end{equation}
of order $K-1$,
where the parameter $\theta \in \R^K$ is a vector of Chebyshev coefficients and
$T_k(\tilde{\Lambda}) \in \R^{n \times n}$ is the Chebyshev polynomial of order
$k$ evaluated at $\tilde{\Lambda} = 2 \Lambda / \lambda_{max} - I_n$, a diagonal
matrix of scaled eigenvalues that lie in $[-1,1]$.
The filtering operation can then be written as $y = g_\theta(L) x
= \sum_{k=0}^{K-1} \theta_k T_k(\tilde{L}) x$, where $T_k(\tilde{L}) \in \R^{n
\times n}$ is the Chebyshev polynomial of order $k$ evaluated at the scaled
Laplacian $\tilde{L} = 2 L / \lambda_{max} - I_n$. %Note that the spectrum of the
%normalized Laplacian is bounded by $2$ \cite{book:Chung97Spectral}, such that
%the scaling can simply be $\tilde{L} = L - I_n$. 
Denoting $\bar{x}_k =
T_k(\tilde{L})x \in \R^n$, we can use the recurrence relation to compute
$\bar{x}_k = 2\tilde{L} \bar{x}_{k-1} - \bar{x}_{k-2}$ with $\bar{x}_0 = x$ and
$\bar{x}_1 = \tilde{L}x$. The entire filtering operation
	$y = g_\theta(L) x = [\bar{x}_0, \ldots, \bar{x}_{K-1}] \theta$
then costs $\bO(K|\E|)$ operations.

@@

Graclus \cite{art:DhillonGuanKulis07Graclus}, built on Metis
\cite{art:KarypisKumar98Metis}, uses a greedy algorithm to compute successive
coarser versions of a given graph and is able to minimize several popular
spectral clustering objectives, from which we chose the normalized cut
\cite{art:ShiMalik00NCut}. Graclus'
greedy rule consists, at each coarsening level, in picking an unmarked
vertex $i$ and matching it with one of its unmarked neighbors $j$ that
maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$. The two matched
vertices are then marked and the coarsened weights are set as the sum of their
weights. The matching is repeated until all nodes have been explored. This is
an very fast coarsening scheme which divides the number of nodes by
approximately two (there may exist a few singletons, non-matched nodes) from
one level to the next coarser level.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{coarsening}
\caption{\textbf{Example of Graph Coarsening and Pooling.} Let us carry out a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
\R^8$ living on $\G_0$, the finest graph given as input. Note that it originally
possesses $n_0 = |\V_0| = 8$ vertices, arbitrarily ordered. For a pooling of
size 4, two coarsenings of size 2 are needed: let Graclus gives $\G_1$ of size
$n_1 = |\V_1| = 5$, then $\G_2$ of size $n_2 = |\V_2| = 3$, the coarsest graph.
Sizes are thus set to $n_2 = 3$, $n_1 = 6$, $n_0 = 12$ and fake nodes (in blue)
are added to $\V_1$ (1 node) and $\V_0$ (4 nodes) to pair with the singeltons
(in orange), such that each node has exactly two children.  Nodes in $\V_2$ are
then arbitrarily ordered and nodes in $\V_1$ and $\V_0$ are ordered
consequently.  At that point the arrangement of vertices in $\V_0$ permits a
regular 1D pooling on $x \in \R^{12}$ such that $z = [\max(x_0,x_1),
\max(x_4,x_5,x_6), \max(x_8,x_9,x_{10})] \in \R^3$, where the signal components
$x_2,x_3,x_7,x_{11}$ are set to a neutral value.}
\label{fig:pooling}
\end{figure}

Pooling operations are carried out many times and must be efficient. After
coarsening, the vertices of the input graph and its coarsened versions are not
arranged in any meaningful way. Hence, a direct application of the pooling
operation would need a table to store all matched vertices. That would result
in a memory inefficient, slow, and hardly parallelizable implementation. It is
however possible to arrange the vertices such that a graph
pooling operation becomes as efficient as a 1D pooling. We proceed in two
steps: (i) create a balanced binary tree and (ii) rearrange the vertices.
After coarsening, each node has either two children, if it was matched at the
finer level, or one, if it was not, i.e the node was a singleton. From the
coarsest to finest level, fake nodes, i.e.  disconnected nodes, are added to
pair with the singletons such that each node has two children. This structure
is a balanced binary tree: (i) regular nodes (and singletons) either have two
regular nodes (e.g. level 1 vertex 0 in \figref{pooling}) or (ii) one singleton
and a fake node as children (e.g. level 2 vertex 0), and
(iii) fake nodes always have two fake nodes as children (e.g. level 1 vertex 1).
Input signals are initialized with a neutral value at the
fake nodes, e.g. 0 when using a ReLU activation with max pooling. Because these
nodes are disconnected, filtering does not impact the initial neutral value.
While those fake nodes do artificially increase the dimensionality thus the
computational cost, we found that, in practice, the number of singletons left
by Graclus is quite low.
% < 10%? compared to the number of vertices.
Arbitrarily ordering the nodes at the coarsest level, then propagating this
ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
children, produces a regular ordering in the finest level. Regular in the sense
that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
rearranged graph signal is analog to pooling a regular 1D signal.
\figref{pooling} shows an example of the whole process.
This regular arrangement makes the operation very efficient and
satisfies parallel architectures such as GPUs as memory accesses are local,
i.e.  matched nodes do not have to be fetched.

@@

The works of \cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF} introduced the
concept of constructing a local receptive field to reduce the number of learned
parameters. The idea is to group together features based upon a measure of
similarity such as to select a limited number of connections between two
successive layers. While this model reduces the number of parameters by
exploiting the locality assumption, it did not attempt to exploit any
stationarity property, i.e. no weight-sharing strategy.
%This approach is however not able to extract similar groups over the data
%domain, as it is not clear how to define convolutional filters on these
%groups.
The authors of \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} used this idea for
their spatial formulation of graph CNNs. They use a weighted graph to define
the local neighborhood and compute a multiscale clustering of the graph for the
pooling operation. Inducing weight sharing in a spatial construction is however
challenging, as it requires to select and order the neighborhoods when a
problem-specific ordering (spatial, temporal, or otherwise) is missing.

A spatial generalization of CNNs to 3D-meshes, a class of smooth
low-dimensional non-Euclidean spaces, was proposed in \cite{masci2015geodesic}.
The authors used geodesic polar coordinates to define the convolution on
mesh patches, and formulated a deep learning architecture which allows
comparison across different manifolds. They obtained state-of-the-art results
for 3D shape recognition.

@@

To validate our model, we applied it to the Euclidean case on the benchmark
MNIST classification problem \cite{pro:LeCunBottouBengioHaffner98MNIST}, a
dataset of 70,000 digits represented on a 2D grid of size $28 \times 28$. For
our graph model, we construct an $8$-NN graph of the 2D grid which produces a
graph of $n = |\V| = 976$ nodes ($28^2 = 784$ pixels and 192 fake nodes as
explained in \secref{pooling}) and $|\E| = 3198$ edges. Following standard
practice, the weights of a $k$-NN similarity graph (between features) are
computed as
%$W_{ij} = \exp(-\|z_i - z_j\|_2^2 / \sigma^2)$
\begin{equation} \label{eq:knngraph}
	W_{ij} = \exp \left( -\frac{\|z_i - z_j\|_2^2}{\sigma^2} \right),
\end{equation}
where $z_i$ is the 2D coordinate of pixel $i$.
% and \sigma is set using the scheme proposed in
